BERT Model Compression and Weight Sharing

### 2019-05-13 paper/paper.tex
Minor fixes

In this paper, we demonstrate that the DeepTwist framework
\citep{lee2018deeptwist} is applicable to compressing fine-tuned BERT
\citep{devlin2018bert} on multiclass classification problems in addition to
computer vision tasks, though the drop in performance is less negligible.
Knowledge distillation also achieves the compression objective with more
sacrifices in  predictive accuracy.

@@

Pruning, quantization, and rank reduction are operations applied to a
particular network's weights without altering the network architecture. 
A different framework, knowledge distillation, first proposed by 
\citep{bucilua2006model} as
``knowledge transfer,'' was recently popularized by
\citep{hinton2015distilling}. In knowledge distillation, a simple network,
called the \emph{student network}, is not trained on the actual data but on
the output of the large network one wishes to compress, called the 
\emph{teacher network}. This procedure can even be extended by inserting
various
intermediate-sized networks between the teacher and student, intuitively
termed
\emph{teacher-assistant networks} \citep{mirzadeh2019improved}. We dicuss
knowledge distillation in detail in \Cref{sec:training}.

@@

In this paper, we limit our purview to text classification tasks. Training
neural-network-based systems for this task typically consists of
gradient-descent-based optimizers applied to a differentiable loss
function, which takes in the prediction for a given example and the correct label. That is, if we have
training data $(\bm{x}_i, y_i)$ and model $f$, we would aim to minimize,
\[ \ell(f(\bm{x_i}), y_i)\]
where usually the loss function is taken to be cross-entropy loss. When we
already have a model $f$ that performs well, we can leverage this training
paradigm through a modification called \emph{knowledge distillation}. Given
the large model that performs well, termed the \emph{teacher network}, we
train a smaller network $g$, the \emph{student network}, by minimizing,
\[ \ell(g(\bm{x_i}), f(\bm{x_i})).\]
We can take
$\ell$ to be either mean-square error (MSE) between the probabilities in
log-space, or Kullback-Leibler (KL) loss. We avoid MSE error on the
probabilities themselves because of disappearing gradients when the predicted
probabilities are very close to 0 or 1.

@@

We have significant flexibility in picking a distortion function, but in order
to be useful, it must provide compression benefits. That is, the image of the
distortion function should have a parameterization that takes less space in
memory than then range of all possible $\btheta$s. For example, if $\btheta \in
\mathbb{R}^{m,n}$, and the distortion produces a matrix of rank 1, then any
distorted matrix could be uniquely identified by $\btheta' \in \mathbb{R}^{n+m}$
through a matrix factorization.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=.5\textwidth]{../figs/value_by_layer.pdf}
\caption{Weight distributions for five most/least variable layers}
\label{fig:weightsbylayer}
\end{figure}

@@

Remarkably, the fact that pruning and rank reduction incurs significant
losses to the
model seems to show that popular compression techniques that enjoy success
in convolutional neural networks fail to generalize good performance to
transformer-based language networks like BERT. We present some preliminary
evidence in \Cref{fig:singular-values}, where we plot the singular values
in one of the matrices in BERT. The effective rank of the matrix seems
quite high, as the singular values are all sizable, decaying at a rate too
slow for rank reduction techniques to be promising. This suggests that
matrices in BERT are in fact effectively dense, unlike CNNs, where
good compression relies on the effective sparsity of the matrices. 
\begin{figure}[tb]
    \centering
    \includegraphics[width=0.5\textwidth]
    {../figs/svd-singular-vals-distribution.png}
    \caption{Distribution of singular values in one of BERT layers}
    \label{fig:singular-values}
\end{figure}

@@

We also identify some curious behavior in the KD training curves. When training
the LSTM on the BERT outputs, neither training loss nor validation loss decrease
at first. In fact, they stay approximately constant for a couple of epochs. Only
after this prolonged period does loss abruptly start going down, producing a
normal curve from that point forward. We find this---depicted in
\cref{fig:kdlstmloss}---remarkably puzzling. We hypothesize this may be the
effect of our adaptive learning rate starting out too low and taking multiple
epochs to grow to a point where it can meaningfully update parameters.

>>>>

In this paper, we demonstrate that the DeepTwist framework
\citep{lee2018deeptwist} is applicable to compressing fine-tuned BERT
\citep{devlin2018bert} on multiclass classification problems in addition to
computer vision tasks, though the drop in performance is less negligible.
Knowledge Distillation also achieves the compression objective with more
sacrifices in  predictive accuracy.

@@

Pruning, quantization, and rank reduction are operations applied to a
particular network's weights without altering the network architecture. 
A different framework, knowledge distillation, first proposed by 
\citet{bucilua2006model} as
``knowledge transfer,'' was recently popularized by
\citet{hinton2015distilling}. In knowledge distillation, a simple network,
called the \emph{student network}, is not trained on the actual data but on
the output of the large network one wishes to compress, called the 
\emph{teacher network}. This procedure can even be extended by inserting
various
intermediate-sized networks between the teacher and student, intuitively
termed
\emph{teacher-assistant networks} \citep{mirzadeh2019improved}. We dicuss
knowledge distillation in detail in \Cref{sec:training}.

@@

In this paper, we limit our purview to text classification tasks. Training
neural-network-based systems for this task typically consists of
gradient-descent-based optimizers applied to a differentiable loss
function, which takes in the prediction for a given example and the correct label. That is, if we have
training data $(\bm{x}_i, y_i)$ and model $f$, we would aim to minimize,
\[ \ell(f(\bm{x_i}), y_i)\]
where usually the loss function is taken to be cross-entropy loss. When we
already have a model $f$ that performs well, we can leverage this training
paradigm through a modification called \emph{knowledge distillation}. Given
the large model that performs well, termed the \emph{teacher network}, we
train a smaller network $g$, the \emph{student network}, by minimizing
\[ \ell(g(\bm{x_i}), f(\bm{x_i})).\]
We can take
$\ell$ to be either mean-square error (MSE) between the probabilities in
log-space, or Kullback-Leibler (KL) loss. We avoid MSE error on the
probabilities themselves because of disappearing gradients when the predicted
probabilities are very close to 0 or 1.

@@

We have significant flexibility in picking a distortion function, but in order
to be useful, it must provide compression benefits. That is, the image of the
distortion function should have a parameterization that takes less space in
memory than then range of all possible $\btheta$s. For example, if $\btheta \in
\mathbb{R}^{m \times n}$, and the distortion produces a matrix of rank 1, then
any
distorted matrix could be uniquely identified by $\btheta' \in \mathbb{R}^{m+n}$
through a matrix factorization.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=.48\textwidth]{../figs/value_by_layer.pdf}
\caption{Weight distributions for five most/least variable layers}
\label{fig:weightsbylayer}
\end{figure}

@@

Remarkably, the fact that pruning and rank reduction incurs significant
losses to the
model seems to show that popular compression techniques that enjoy success
in convolutional neural networks fail to generalize good performance to
transformer-based language networks like BERT. We present some preliminary
evidence in \Cref{fig:singular-values}, where we plot the singular values
in one of the matrices in BERT. The effective rank of the matrix seems
quite high, as the singular values are all sizable, decaying at a rate too
slow for rank reduction techniques to be promising. This suggests that
matrices in BERT are in fact effectively dense, unlike CNNs, where
good compression relies on the effective sparsity of the matrices. 
\begin{figure}[tb]
    \centering
    \includegraphics[width=0.46\textwidth]
    {../figs/svd-singular-vals-distribution.png}
    \caption{Distribution of singular values in one of BERT layers}
    \label{fig:singular-values}
\end{figure}

@@

We also identify some curious behavior in the KD training curves. When training
the LSTM on the BERT outputs, neither training loss nor validation loss decrease
at first. In fact, they stay approximately constant for a couple of epochs. Only
after this prolonged period does loss abruptly start going down, producing a
normal curve from that point forward. We find this---depicted in
\cref{fig:kdlstmloss}---remarkably puzzling. We speculate this may be the
effect of our adaptive learning rate starting out too low and taking multiple
epochs to grow to a point where it can meaningfully update parameters.
